---
always_allow_html: yes
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Introduction to NOAAeq

The `NOAAeq` package analyses data from the [Significant Earthquake Database](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) of the U.S. National Oceanographic and Atmospheric Administration (NOAA). This dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.

The package has 3 main functionalities:

* Data cleaning

* Timeline plots

* Interactive leaflet maps

## Timelines

Here is an example of a timeline plot for the USA, covering the recent period 2000-2017.

```{r setup, include=FALSE}
require("knitr")
require("NOAAeq")
require("magrittr")
opts_knit$set(root.dir = system.file("extdata", package = "NOAAeq"))
```

```{r data, include=FALSE}
data <- readr::read_delim(file = "earthquakes.tsv.gz", delim = "\t")
```

```{r timeline1, echo=TRUE, fig.width=7, fig.height=2}
data %>%
 eq_clean_data() %>%
 dplyr::filter(lubridate::year(DATE) %in% 2000:2017 & COUNTRY == "USA") %>%
 ggplot(aes(x = DATE, size = EQ_PRIMARY, fill = TOTAL_DEATHS)) +
 geom_timeline() +
 theme_classic() +
 theme(legend.position = "bottom") +
 scale_size_continuous(name = "Richter scale value") +
 scale_fill_continuous(name = "# deaths") +
 guides(size = guide_legend(order = 1),
        fill = guide_colourbar(order = 2))
```

[![Travis-CI Build Status](https://travis-ci.org/la-sch/NOAAeq.svg?branch=master)](https://travis-ci.org/la-sch/NOAAeq)
